---
title: "hictkR Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{hictkR Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(hictkR)
```

## Opening files

hictkR supports opening .hic and .cool files using the `File()` function.

When opening files in .hic format, `resolution` is a mandatory parameter.

```{r}
path <- system.file("extdata", "interactions.hic", package="hictkR")
f <- File(path, resolution=100000)
```

hictkR file handles returned by `File()` provide several attributes storing static properties of the opened file

```{r, results="hide"}
path <- f$path
resolution <- as.integer(f$bin_size)

paste("File \"",
      basename(path),
      "\" has the following resolution: ",
      resolution,
      "bp",
      sep="")
```

__File "`r basename(path)`" has the following resolution: `r resolution` bp__


## Accessing chromosomes and bins

Some of these properties store the list of chromosomes as well as the bin table associated with the opened file.

```{r}
chroms <- f$chromosomes
bins <- f$bins
normalizations <- f$normalizations
```

```{r, echo=FALSE}
chroms_ <- t(chroms)
knitr::kable(chroms_, format="html", caption="Chromosomes")
```

```{r, echo=FALSE}
bins_ <- head(bins)

bins_[] <- lapply(bins_, as.character)
bins_[nrow(bins_) + 1,] <- "..."

knitr::kable(t(bins_), format="html", caption="Bin table", col.names=NULL)
```


```{r, echo=FALSE}
normalizations_ <- t(normalizations)
knitr::kable(normalizations_, format="html", caption="Normalizations", col.names=NULL)
```

## Fetching interactions

hictkR support fetching interactions through the `fetch()` function.

#### Fetch genome-wide interactions with genomic coordinates

```{r, results="hide", eval=FALSE}
fetch(f, join=TRUE)
```
```{r, echo=FALSE}
df <- head(fetch(f, join=TRUE))
knitr::kable(df, format="html", format.args=list(scientific=FALSE))
```

Interactions can also be normalized using one of the available normalization methods (see `f$normalizations`).

```{r, results="hide", eval=FALSE}
fetch(f, join=TRUE, normalization="ICE")
```
```{r, echo=FALSE}
df <- head(fetch(f, join=TRUE, normalization="ICE"))
knitr::kable(df, format="html", format.args=list(scientific=FALSE))
```

#### Fetch interactions for a region of interest

hictkR can be instructed to fetch interactions for a region of interest.

This is much more efficient than fetching genome-wide interactions and then filtering interactions using R.

```{r, results="hide", eval=FALSE}
fetch(f, "chr2L", join=TRUE)
```
```{r, echo=FALSE}
df <- head(fetch(f, "chr2L", join=TRUE))
knitr::kable(df, format="html", format.args=list(scientific=FALSE))
```

```{r, results="hide", eval=FALSE}
fetch(f,
      "chr2L:0-10,000,000",
      "chr3R:10,000,000-20,000,000",
      join=TRUE)
```
```{r, echo=FALSE}
df <- head(fetch(f, "chr2L:0-10,000,000", "chr3R:10,000,000-20,000,000", join=TRUE))
knitr::kable(df, format="html", format.args=list(scientific=FALSE))
```

#### Fetch interactions as a dense matrix

hictkR can return interactions as a `Matrix` when specifying `type="dense"`.

```{r}
fetch(f,
      "chr2L:0-1,000,000",
      "chr3R:1,000,000-2,000,000",
      type="dense")
```

This can be very useful to visualize interaction matrices as heatmaps.

```{r, show_figure, fig.width = 7.5, fig.height = 7.5}
m <- fetch(f, "chr3R:0-20,000,000", normalization="ICE", type="dense")
image(log(m), axes=FALSE)
```
